---
title: "CLONEID primary use case figure reproduction"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  subtree_rds: "input/subtree/subtree_SNU-668_dp_A9_seedT2.rds"
  panel_a_zip: "input/subtree/panelA.zip"
  panel_de_zip: "input/subtree/panelDE.zip"
  manifest_a_yaml: "input/manifest_A/manifest_panel_A.yaml"
  manifest_de_yaml: "input/manifest_DE/manifest_panel_DE.yaml"
  output_panels_dir: "output/panels"
  output_final_dir: "output/final"
  cache_dir: "cache"
  root_dir: "."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r init}
source(file.path(params$root_dir, "R", "io.R"))
source(file.path(params$root_dir, "R", "panel_bc.R"))
source(file.path(params$root_dir, "R", "panel_ade.R"))

paths <- cloneid_figure_paths(
  root_dir = params$root_dir,
  subtree_rds = params$subtree_rds,
  panel_a_zip = params$panel_a_zip,
  panel_de_zip = params$panel_de_zip,
  manifest_a_yaml = params$manifest_a_yaml,
  manifest_de_yaml = params$manifest_de_yaml,
  output_panels_dir = params$output_panels_dir,
  output_final_dir = params$output_final_dir,
  cache_dir = params$cache_dir
)

status <- describe_primary_figure_inputs(paths)
status
```

```{r derive-manifests}
panel_a_manifest <- derive_panel_a_manifest(
  subtree_rds = paths$subtree_rds,
  output_yaml = paths$manifest_a_yaml
)

panel_de_manifest <- derive_panel_de_manifest(
  subtree_rds = paths$subtree_rds,
  output_yaml = paths$manifest_de_yaml
)

status <- describe_primary_figure_inputs(paths)
```

# Scope

This workflow is intended to reproduce the manuscript figure:

```
CLONEID links event history, growth dynamics and karyotypic state in a longitudinal adaptation experiment.
```

The scaffold is organized around the approved panel split:

- `A`: circular phylogeny rooted at `SNU-668_K3_A24_seedT4`, with one first-available cropped overlay per node from `panelA.zip`
- `B`: daily cell counts and fitted growth curves for `r`- and `K`-selected replicate lineages
- `C1-C4`: trend lines for growth rate and observed maximum population size, split by `r` versus `K`
- `D/E`: late genome-derived summaries from the whole-subtree manifest export

# Input status

```{r input-status}
knitr::kable(status, align = "ll")
```

# Manual portal steps

Render this document first. It will regenerate the two manifest YAML files from the subtree bundle and place them in the configured input directories.

```{r manual-steps}
knitr::kable(manual_portal_steps(paths), align = "ll")
```

The intended manual sequence is:

1. Ensure `input/subtree/subtree_SNU-668_dp_A9_seedT2.rds` is present.
2. Render this `.Rmd` to regenerate:
   - `input/manifest_A/manifest_panel_A.yaml`
   - `input/manifest_DE/manifest_panel_DE.yaml`
3. Upload `manifest_panel_A.yaml` in the portal and copy the resulting `panelA.zip` into `input/subtree/`.
4. Upload `manifest_panel_DE.yaml` in the portal and copy the resulting `panelDE.zip` into `input/subtree/`.
5. Re-render the `.Rmd` once `panelA.zip` and `panelDE.zip` are present.

# Downloaded archive verification

If manifest-driven downloads have been copied into `input/subtree/` as `panel*.zip`, this block checks whether their contents match the expected panel targets.

```{r panel-zip-verification}
panel_zip_check <- verify_panel_zip_contents(
  zip_dir = file.path(params$root_dir, "input", "subtree"),
  panel_a_manifest = panel_a_manifest,
  panel_de_manifest = panel_de_manifest
)

knitr::kable(panel_zip_check$summary, align = "llllll")
```

```{r panel-zip-missing-targets}
if (length(panel_zip_check$details) > 0) {
  for (detail in panel_zip_check$details) {
    cat("\n\n###", basename(detail$zip_file), "\n\n")
    cat("Detected panel:", detail$detected_panel, "\n\n")
    if (length(detail$missing_targets) == 0) {
      cat("All expected targets were detected.\n")
    } else {
      print(detail$missing_targets)
    }
  }
} else {
  cat("No panel*.zip files found under input/subtree/.\n")
}
```

# Panel B/C data inspection

```{r subtree-summary}
subtree_summary <- summarize_subtree_for_growth(paths$subtree_rds)
subtree_summary$bundle_overview
```

```{r bc-ready-check}
bc_inputs <- inspect_bc_inputs(paths$subtree_rds)
bc_inputs$summary
```

# Panel A

```{r panel-a0-preview, fig.width=4, fig.height=12}
plot_panel_a0(paths$subtree_rds)
```

```{r panel-a-manifest-preview}
knitr::kable(utils::head(panel_a_manifest, 12), align = "ll")
```

```{r panel-a-status}
panel_a_status(paths$manifest_a_yaml, paths$panel_a_zip)
```

```{r panel-a-render}
panel_a_outputs <- tryCatch(
  save_panel_a_outputs(
    subtree_rds = paths$subtree_rds,
    root_dir = params$root_dir,
    panel_a_zip = paths$panel_a_zip,
    output_panels_dir = paths$output_panels_dir,
    cache_dir = paths$cache_dir
  ),
  error = function(e) e
)

if (inherits(panel_a_outputs, "error")) {
  cat("Panel A rendering not available yet:", conditionMessage(panel_a_outputs), "\n")
} else {
  knitr::kable(panel_a_outputs$files, align = "ll")
}
```

```{r panel-a-preview, fig.width=10, fig.height=10}
if (!inherits(panel_a_outputs, "error")) {
  plot_panel_a(
    subtree_rds = paths$subtree_rds,
    panel_a_zip = paths$panel_a_zip,
    cache_dir = paths$cache_dir
  )
}
```

```{r panel-a-overlays}
if (!inherits(panel_a_outputs, "error")) {
  knitr::kable(utils::head(panel_a_outputs$overlay_map, 12), align = "ll")
}
```

# Panel B

```{r panel-b-render}
bc_outputs <- save_bc_outputs(
  subtree_rds = paths$subtree_rds,
  output_panels_dir = paths$output_panels_dir,
  cache_dir = paths$cache_dir
)

panel_b_status()
```

# Panel C

```{r panel-b-selected-passages}
knitr::kable(bc_outputs$panel_b_data$selected_passages, align = "ll")
```

```{r panel-b-artifacts}
knitr::kable(bc_outputs$files, align = "ll")
```

```{r panel-b-preview, fig.width=14, fig.height=6.5}
plot_panel_b(bc_outputs$panel_b_data)
```

```{r panel-c-placeholder}
panel_c_status()
```

```{r panel-c-preview, fig.width=12, fig.height=10}
draw_panel_c_combined(bc_outputs$panel_c_plots)
```

```{r panel-c-metrics-head}
knitr::kable(utils::head(bc_outputs$metrics, 12), align = "lllllll")
```

# Panel D/E placeholder

```{r panel-de-manifest-preview}
knitr::kable(panel_de_manifest, align = "ll")
```

```{r panel-de-placeholder}
panel_de_status(paths$manifest_de_yaml, paths$panel_de_zip)
```

```{r panel-de-render}
de_outputs <- tryCatch(
  save_de_outputs(
    root_dir = params$root_dir,
    panel_de_zip = paths$panel_de_zip,
    output_panels_dir = paths$output_panels_dir,
    cache_dir = paths$cache_dir
  ),
  error = function(e) e
)

if (inherits(de_outputs, "error")) {
  print(de_outputs)
  cat("Panel D/E rendering not available yet:", conditionMessage(de_outputs), "\n")
} else {
  knitr::kable(de_outputs$files, align = "ll")
}
```

```{r panel-d-preview, fig.width=8, fig.height=5}
if (!inherits(de_outputs, "error")) {
  de_outputs$panel_d_plot
}
```

```{r panel-e-preview, fig.width=12, fig.height=5.5}
if (!inherits(de_outputs, "error")) {
  grid::grid.newpage()
  grid::grid.draw(de_outputs$panel_e_plot$gtable)
}
```

# Final assembly

```{r final-assembly-render}
final_outputs <- tryCatch(
  assemble_primary_use_case_figure(
    output_panels_dir = paths$output_panels_dir,
    output_final_dir = paths$output_final_dir,
    subtree_rds = paths$subtree_rds,
    panel_a_zip = paths$panel_a_zip,
    cache_dir = paths$cache_dir
  ),
  error = function(e) e
)

if (inherits(final_outputs, "error")) {
  cat("Final assembly not available yet:", conditionMessage(final_outputs), "\n")
} else {
  knitr::kable(final_outputs$files, align = "ll")
}
```

```{r final-assembly-status}
final_assembly_status(paths$output_panels_dir, paths$output_final_dir)
```

```{r final-assembly-preview, fig.width=16, fig.height=22}
if (!inherits(final_outputs, "error")) {
  knitr::include_graphics(final_outputs$output_png)
}
```
