---
title: "CLONEID primary use case figure reproduction"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  subtree_rds: "input/subtree/subtree_SNU-668_dp_A9_seedT2.rds"
  panel_a_zip: "input/subtree/panelA.zip"
  panel_de_zip: "input/subtree/panelDE.zip"
  manifest_a_yaml: "input/manifest_A/manifest_panel_A.yaml"
  manifest_de_yaml: "input/manifest_DE/manifest_panel_DE.yaml"
  output_panels_dir: "output/panels"
  output_final_dir: "output/final"
  cache_dir: "cache"
  root_dir: "."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
required_packages <- c(
  "yaml", "jsonlite", "ggplot2", "pheatmap", 
  "scales", "wesanderson", "magick", "png", "ape"
)
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) {
  install.packages(new_packages, repos = "[http://cran.us.r-project.org](http://cran.us.r-project.org)")
}
```

```{r init}
source(file.path(params$root_dir, "R", "io.R"))
source(file.path(params$root_dir, "R", "panel_bc.R"))
source(file.path(params$root_dir, "R", "panel_ade.R"))

paths <- cloneid_figure_paths(
  root_dir = params$root_dir,
  subtree_rds = params$subtree_rds,
  panel_a_zip = params$panel_a_zip,
  panel_de_zip = params$panel_de_zip,
  manifest_a_yaml = params$manifest_a_yaml,
  manifest_de_yaml = params$manifest_de_yaml,
  output_panels_dir = params$output_panels_dir,
  output_final_dir = params$output_final_dir,
  cache_dir = params$cache_dir
)

status <- describe_primary_figure_inputs(paths)
status
```

```{r derive-manifests}
if (file.exists(paths$subtree_rds)) {
  panel_b_manifest <- derive_panel_b_manifest(
    subtree_rds = paths$subtree_rds,
    output_yaml = paths$manifest_a_yaml
  )

  panel_de_manifest <- derive_panel_de_manifest(
    subtree_rds = paths$subtree_rds,
    output_yaml = paths$manifest_de_yaml
  )

  status <- describe_primary_figure_inputs(paths)
} else {
  panel_b_manifest <- data.frame(
    note = "Manifest generation skipped: missing subtree RDS bundle.",
    stringsAsFactors = FALSE
  )
  panel_de_manifest <- data.frame(
    note = "Manifest generation skipped: missing subtree RDS bundle.",
    stringsAsFactors = FALSE
  )
  cat(
    "Missing required input:", paths$subtree_rds, "\n",
    "Download the subtree .rds bundle from dev.cloneid.org first, then re-render to generate the panel B and panel E/F manifests.\n"
  )
}
```

# Scope

This workflow is intended to reproduce the manuscript figure:

```         
CLONEID links event history, growth dynamics and karyotypic state in a longitudinal adaptation experiment.
```

The scaffold is organized around the approved panel split:

-   `A`: event-history overview for the full subtree, highlighting the lineage used in panel `B`
-   `B`: circular phylogeny rooted at `SNU-668_K3_A24_seedT4`, with one first-available cropped overlay per node from `panelA.zip`
-   `C`: daily cell counts and fitted growth curves for `r`- and `K`-selected replicate lineages
-   `D1-D4`: trend lines for growth rate and observed maximum population size, split by `r` versus `K`
-   `E/F`: late genome-derived summaries from the whole-subtree manifest export

# Input status

```{r input-status}
knitr::kable(status, align = "ll")
```

```{r missing-input-notices, results='asis'}
missing_inputs <- character()
if (!file.exists(paths$subtree_rds)) {
  missing_inputs <- c(
    missing_inputs,
    sprintf("- Missing subtree bundle: `%s`", paths$subtree_rds)
  )
}
if (!file.exists(paths$panel_a_zip)) {
  missing_inputs <- c(
    missing_inputs,
    sprintf("- Missing Panel B archive source zip: `%s`", paths$panel_a_zip)
  )
}
if (!file.exists(paths$panel_de_zip)) {
  missing_inputs <- c(
    missing_inputs,
    sprintf("- Missing Panel E/F archive: `%s`", paths$panel_de_zip)
  )
}

if (length(missing_inputs) == 0) {
  cat("All required input files are present.\n")
} else {
  cat("## Missing required inputs\n\n")
  cat(paste(missing_inputs, collapse = "\n"), "\n\n")
  cat("Render can still proceed to generate status tables and any manifests allowed by the available inputs, but panel rendering will be skipped until the missing files are supplied.\n")
}
```

# Manual portal steps

Render this document first. It will regenerate the two manifest YAML files from the subtree bundle and place them in the configured input directories.

```{r manual-steps}
knitr::kable(manual_portal_steps(paths), align = "ll")
```

The intended manual sequence is:

1.  Ensure `input/subtree/subtree_SNU-668_dp_A9_seedT2.rds` is present. 
    Here is a brief steps to obtain this subtree file: 
      1) open https://dev.cloneid.org/cloneid, then find module "View content of CLONEID DB" and click the `open` bottom;
      2) paste `SNU-668_dp_A9_seedT2` into the search box;
      3) Then, on the page that opens, click the root node `SNU-668_dp_A9_seedT2`, and in the pop-up submenu, click `Download subtree (RDS)`;
      4) copy `subtree_SNU-668_dp_A9_seedT2.rds` into `input/subtree/`.
2.  Render this `.Rmd` to regenerate:
    -   `input/manifest_A/manifest_panel_A.yaml`
    -   `input/manifest_DE/manifest_panel_DE.yaml`
3.  Upload `manifest_panel_A.yaml` to <https://dev.cloneid.org/manifest-download>. After the job finished, click `Download Package` to download `package.zip`. Then rename it to `panelA.zip`, and copy it into `input/subtree/`.
4.  Upload `manifest_panel_DE.yaml` to <https://dev.cloneid.org/manifest-download> and copy the resulting as `panelDE.zip` into `input/subtree/`.
5.  Re-render the `.Rmd` once `panelA.zip` and `panelDE.zip` are present.

# Downloaded archive verification

If manifest-driven downloads have been copied into `input/subtree/` as `panel*.zip`, this block checks whether their contents match the expected panel targets.

```{r panel-zip-verification}
if (file.exists(paths$subtree_rds)) {
  panel_zip_check <- verify_panel_zip_contents(
    zip_dir = file.path(params$root_dir, "input", "subtree"),
    panel_a_manifest = panel_b_manifest,
    panel_de_manifest = panel_de_manifest
  )

  knitr::kable(panel_zip_check$summary, align = "llllll")
} else {
  panel_zip_check <- list(details = list())
  cat("Zip verification skipped because the subtree .rds bundle is missing and manifests could not be generated.\n")
}
```

```{r panel-zip-missing-targets}
if (length(panel_zip_check$details) > 0) {
  for (detail in panel_zip_check$details) {
    cat("\n\n###", basename(detail$zip_file), "\n\n")
    cat("Detected panel:", detail$detected_panel, "\n\n")
    if (length(detail$missing_targets) == 0) {
      cat("All expected targets were detected.\n")
    } else {
      print(detail$missing_targets)
    }
  }
} else {
  cat("No panel*.zip files found under input/subtree/.\n")
}
```

# Panel C/D data inspection

```{r subtree-summary}
if (file.exists(paths$subtree_rds)) {
  subtree_summary <- summarize_subtree_for_growth(paths$subtree_rds)
  subtree_summary$bundle_overview
} else {
  cat("Panel C/D data inspection skipped because the subtree .rds bundle is missing.\n")
}
```

```{r bc-ready-check}
if (file.exists(paths$subtree_rds)) {
  bc_inputs <- inspect_bc_inputs(paths$subtree_rds)
  bc_inputs$summary
} else {
  cat("Panel C/D readiness check skipped because the subtree .rds bundle is missing.\n")
}
```

# Panel A

```{r panel-a-preview, fig.width=4, fig.height=12}
if (file.exists(paths$subtree_rds)) {
  plot_panel_a(paths$subtree_rds)
} else {
  cat("Panel A preview skipped because the subtree .rds bundle is missing.\n")
}
```

```{r panel-a-render}
panel_a_outputs <- NULL
if (!file.exists(paths$subtree_rds)) {
  cat("Panel A rendering not available yet: missing subtree .rds bundle at", paths$subtree_rds, "\n")
} else {
  panel_a_outputs <- save_panel_a_outputs(
    subtree_rds = paths$subtree_rds,
    output_panels_dir = paths$output_panels_dir
  )
  knitr::kable(panel_a_outputs$files, align = "ll")
}
```

# Panel B

```{r panel-b-manifest-preview}
knitr::kable(utils::head(panel_b_manifest, 12), align = "ll")
```

```{r panel-b-status}
panel_b_status(paths$manifest_a_yaml, paths$panel_a_zip)
```

```{r panel-b-render}
panel_b_outputs <- NULL
if (!file.exists(paths$subtree_rds)) {
  cat("Panel B rendering not available yet: missing subtree .rds bundle at", paths$subtree_rds, "\n")
} else if (!file.exists(paths$panel_a_zip)) {
  cat("Panel B rendering not available yet: missing panelA.zip archive at", paths$panel_a_zip, "\n")
} else {
  panel_b_outputs <- tryCatch(
    save_panel_b_outputs(
      subtree_rds = paths$subtree_rds,
      root_dir = params$root_dir,
      panel_a_zip = paths$panel_a_zip,
      output_panels_dir = paths$output_panels_dir,
      cache_dir = paths$cache_dir
    ),
    error = function(e) e
  )

  if (inherits(panel_b_outputs, "error")) {
    cat("Panel B rendering not available yet:", conditionMessage(panel_b_outputs), "\n")
  } else {
    knitr::kable(panel_b_outputs$files, align = "ll")
  }
}
```

```{r panel-b-preview, fig.width=10, fig.height=10}
if (!is.null(panel_b_outputs) && !inherits(panel_b_outputs, "error")) {
  plot_panel_b(
    subtree_rds = paths$subtree_rds,
    panel_a_zip = paths$panel_a_zip,
    cache_dir = paths$cache_dir
  )
}
```

```{r panel-b-overlays}
if (!is.null(panel_b_outputs) && !inherits(panel_b_outputs, "error")) {
  knitr::kable(utils::head(panel_b_outputs$overlay_map, 12), align = "ll")
}
```

# Panel C

```{r panel-c-render}
cd_outputs <- NULL
if (!file.exists(paths$subtree_rds)) {
  cat("Panel C/D rendering not available yet: missing subtree .rds bundle at", paths$subtree_rds, "\n")
} else {
  cd_outputs <- save_cd_outputs(
    subtree_rds = paths$subtree_rds,
    output_panels_dir = paths$output_panels_dir,
    cache_dir = paths$cache_dir
  )

  panel_c_status()
}
```

# Panel D

```{r panel-c-selected-passages}
if (!is.null(cd_outputs)) {
  knitr::kable(cd_outputs$panel_c_data$selected_passages, align = "ll")
}
```

```{r panel-cd-artifacts}
if (!is.null(cd_outputs)) {
  knitr::kable(cd_outputs$files, align = "ll")
}
```

```{r panel-c-preview, fig.width=14, fig.height=6.5}
if (!is.null(cd_outputs)) {
  plot_panel_c(cd_outputs$panel_c_data)
}
```

```{r panel-d-placeholder}
if (!is.null(cd_outputs)) {
  panel_d_status()
}
```

```{r panel-d-preview, fig.width=12, fig.height=10}
if (!is.null(cd_outputs)) {
  draw_panel_d_combined(cd_outputs$panel_d_plots)
}
```

```{r panel-d-metrics-head}
if (!is.null(cd_outputs)) {
  knitr::kable(utils::head(cd_outputs$metrics, 12), align = "lllllll")
}
```

# Panel E/F placeholder

```{r panel-de-manifest-preview}
knitr::kable(panel_de_manifest, align = "ll")
```

```{r panel-ef-placeholder}
panel_ef_status(paths$manifest_de_yaml, paths$panel_de_zip)
```

```{r panel-ef-render}
ef_outputs <- NULL
if (!file.exists(paths$panel_de_zip)) {
  cat("Panel E/F rendering not available yet: missing panelDE.zip archive at", paths$panel_de_zip, "\n")
} else {
  ef_outputs <- tryCatch(
    save_ef_outputs(
      root_dir = params$root_dir,
      panel_de_zip = paths$panel_de_zip,
      output_panels_dir = paths$output_panels_dir,
      cache_dir = paths$cache_dir
    ),
    error = function(e) e
  )

  if (inherits(ef_outputs, "error")) {
    cat("Panel E/F rendering not available yet:", conditionMessage(ef_outputs), "\n")
  } else {
    knitr::kable(ef_outputs$files, align = "ll")
  }
}
```

```{r panel-e-preview, fig.width=8, fig.height=5}
if (!is.null(ef_outputs) && !inherits(ef_outputs, "error")) {
  ef_outputs$panel_e_plot
}
```

```{r panel-f-preview, fig.width=12, fig.height=5.5}
if (!is.null(ef_outputs) && !inherits(ef_outputs, "error")) {
  grid::grid.newpage()
  grid::grid.draw(ef_outputs$panel_f_plot$gtable)
}
```

# Final assembly

```{r final-assembly-render}
final_outputs <- NULL
if (!file.exists(paths$subtree_rds)) {
  cat("Final assembly not available yet: missing subtree .rds bundle at", paths$subtree_rds, "\n")
} else if (!file.exists(paths$panel_a_zip)) {
  cat("Final assembly not available yet: missing panelB source archive at", paths$panel_a_zip, "\n")
} else if (is.null(cd_outputs)) {
  cat("Final assembly not available yet: panel C/D outputs have not been rendered because the subtree bundle is missing or panel C/D rendering was skipped.\n")
} else if (is.null(ef_outputs) || inherits(ef_outputs, "error")) {
  cat("Final assembly not available yet: panel E/F outputs have not been rendered because panelDE.zip is missing or panel E/F rendering failed.\n")
} else {
  final_outputs <- tryCatch(
    assemble_primary_use_case_figure(
      output_panels_dir = paths$output_panels_dir,
      output_final_dir = paths$output_final_dir,
      subtree_rds = paths$subtree_rds,
      panel_a_zip = paths$panel_a_zip,
      cache_dir = paths$cache_dir
    ),
    error = function(e) e
  )

  if (inherits(final_outputs, "error")) {
    cat("Final assembly not available yet:", conditionMessage(final_outputs), "\n")
  } else {
    knitr::kable(final_outputs$files, align = "ll")
  }
}
```

```{r final-assembly-status}
final_assembly_status(paths$output_panels_dir, paths$output_final_dir)
```

```{r final-assembly-preview, fig.width=16, fig.height=22}
if (!is.null(final_outputs) && !inherits(final_outputs, "error")) {
  knitr::include_graphics(final_outputs$output_png)
}
```
