---
title: "Figure S Use Case GBM Reanalysis"
output:
  html_document:
    toc: true
    toc_float: true
params:
  focus_pid: "P2"
  minobs: 20
  minobs_values: [5, 10, 20]
  n_cores: 9
  pm_values: [5e-05]
  nboot: 45
  n0: 100000
  nb: 10000000
  correct_efflux: true
  force_refit: false
  force_recompute_panels: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  fig.width = 8,
  fig.height = 6
)
```

## Overview

This notebook rebuilds Figure S5 GBM use-case panels B-G from the data stored under
`cloneidR/analysis/FigureS5/data`, while using the current ALFA-K package and the
updated analysis logic used in PANcanKFLs.

Key rules implemented here:

- Patients are discovered from the actual subdirectories currently present under `data/`.
- Metadata come from `meta_data.xlsx`.
- If `ID_alt` is non-missing, it is used in preference to `ID`.
- Sample profiles are read from `data/<pid>/perspectives/<effective_id>/TranscriptomePerspective.rds`.
- The primary-to-recurrent interval follows the PANcanKFLs metadata convention:
  `Primary = 1`, `Recurrent = abs(Age_recurrent - Age_primary) * 360` days.
- ALFA-K fitting is re-run from these inputs across `minobs = 5, 10, 20`,
  with task-level parallelization over `patient x minobs`, and the best
  cross-validated fit is selected separately for each patient.

```{r paths-and-libraries}
suppressPackageStartupMessages({
  library(alfakR)
  library(circlize)
  library(ComplexHeatmap)
  library(cowplot)
  library(dplyr)
  library(doParallel)
  library(foreach)
  library(ggplot2)
  library(ggpubr)
  library(grid)
  library(readxl)
  library(RColorBrewer)
  library(scales)
  library(stringr)
  library(tibble)
  library(tidyr)
  library(uwot)
})

analysis_dir <- "/Users/4482173/Documents/GitHub/cloneidR/analysis/FigureS5"
data_dir <- file.path(analysis_dir, "data")
meta_path <- file.path(data_dir, "meta_data.xlsx")

pancan_repo <- "/Users/4482173/Documents/GitHub/PANcanKFLs"
usecase_repo <- "/Users/4482173/Documents/GitHub/CLONEID_usecases/usecase3"

generated_dir <- file.path(analysis_dir, "generated")
result_dir <- file.path(generated_dir, "results")
figure_dir <- file.path(generated_dir, "figures")
input_dir <- file.path(result_dir, "ALFA-K_inputs")
fit_dir <- file.path(result_dir, "ALFAK_fitnessLandscape")
cache_dir <- file.path(generated_dir, "cache")

dir.create(generated_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(result_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(fit_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)

focus_pid <- params$focus_pid
minobs_values_use <- params$minobs_values
if (is.null(minobs_values_use) || !length(unlist(minobs_values_use))) {
  minobs_values_use <- params$minobs
}
minobs_values_use <- sort(unique(as.integer(unlist(minobs_values_use))))
minobs_values_use <- minobs_values_use[is.finite(minobs_values_use) & minobs_values_use > 0]
if (!length(minobs_values_use)) {
  stop("No valid minobs values supplied.")
}
pm_values <- as.numeric(unlist(params$pm_values))
n_cores_use <- suppressWarnings(as.integer(params$n_cores))
if (!is.finite(n_cores_use) || n_cores_use < 1L) {
  detected_cores <- suppressWarnings(parallel::detectCores(logical = FALSE))
  if (!is.finite(detected_cores) || detected_cores < 1L) {
    detected_cores <- 1L
  }
  n_cores_use <- max(1L, detected_cores - 1L)
}
nboot_use <- as.integer(params$nboot)
n0_use <- as.numeric(params$n0)
nb_use <- as.numeric(params$nb)
correct_efflux_use <- isTRUE(params$correct_efflux)
force_refit <- isTRUE(params$force_refit)
force_recompute_panels <- isTRUE(params$force_recompute_panels)

stage_levels <- c("Primary", "Recurrent")
diploid_state <- paste(rep(2, 22), collapse = ".")
set.seed(3.1415926)
```

## Helper Functions

```{r helpers}
sort_pid_levels <- function(x) {
  x <- unique(as.character(x))
  ord_num <- suppressWarnings(as.integer(sub("^P", "", x)))
  x[order(ifelse(is.na(ord_num), Inf, ord_num), x)]
}

normalize_01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  if (!all(is.finite(rng)) || diff(rng) == 0) {
    return(rep(0, length(x)))
  }
  (x - rng[1]) / diff(rng)
}

normalize_mean <- function(x) {
  rng <- range(x, na.rm = TRUE)
  if (!all(is.finite(rng)) || diff(rng) == 0) {
    return(rep(0, length(x)))
  }
  (x - mean(x, na.rm = TRUE)) / diff(rng)
}

normalize_median <- function(x) {
  rng <- range(x, na.rm = TRUE)
  if (!all(is.finite(rng)) || diff(rng) == 0) {
    return(rep(0, length(x)))
  }
  (x - stats::median(x, na.rm = TRUE)) / diff(rng)
}

gen_all_neighbours <- function(ids, as.strings = TRUE) {
  if (as.strings) {
    ids <- lapply(ids, function(xx) as.numeric(unlist(strsplit(xx, split = "[.]"))))
  }
  nkern <- do.call(rbind, lapply(seq_len(length(ids[[1]])), function(i) {
    x0 <- rep(0, length(ids[[1]]))
    x1 <- x0
    x0[i] <- -1
    x1[i] <- 1
    rbind(x0, x1)
  }))
  n <- do.call(rbind, lapply(ids, function(xx) {
    t(apply(nkern, 1, function(kk) kk + xx))
  }))
  n <- unique(n)
  nids <- length(ids)
  n <- rbind(do.call(rbind, ids), n)
  n <- unique(n)
  n[-seq_len(nids), , drop = FALSE]
}

ensure_center_in_limits <- function(limits, center = 1) {
  c(min(limits[1], center), max(limits[2], center))
}

placeholder_plot <- function(label) {
  ggplot() +
    annotate("text", x = 0, y = 0, label = label, size = 5) +
    theme_void()
}

save_plot_pdf <- function(plot_obj, file, width = 5.5, height = 5) {
  grDevices::pdf(file, width = width, height = height, useDingbats = FALSE)
  on.exit(grDevices::dev.off(), add = TRUE)
  print(plot_obj)
}

read_transcriptome_perspective <- function(rds_path) {
  obj <- readRDS(rds_path)
  stopifnot(is.list(obj), "profile" %in% names(obj))
  profile <- as.matrix(obj$profile)
  storage.mode(profile) <- "numeric"
  list(
    passaging_id = obj$passaging_id,
    perspective_type = obj$perspective_type,
    profile = round(profile)
  )
}

compute_delta_time_days <- function(df_pid) {
  df_pid <- df_pid %>%
    mutate(Stage = factor(Stage, levels = stage_levels)) %>%
    arrange(Stage, effective_id)

  age_primary <- suppressWarnings(as.numeric(df_pid$Age[df_pid$Stage == "Primary"][1]))
  age_recurrent <- suppressWarnings(as.numeric(df_pid$Age[df_pid$Stage == "Recurrent"][1]))

  recurrent_days <- if (is.finite(age_primary) && is.finite(age_recurrent)) {
    as.integer(round(abs(age_recurrent - age_primary) * 360))
  } else {
    NA_integer_
  }
  if (!is.finite(recurrent_days) || recurrent_days <= 1L) {
    recurrent_days <- 180L
  }

  df_pid %>%
    mutate(
      Delta_Time = ifelse(as.character(Stage) == "Primary", 1L, recurrent_days)
    )
}

available_pid_dirs <- function(base_dir) {
  dd <- list.dirs(base_dir, recursive = FALSE, full.names = TRUE)
  dd <- dd[file.info(dd)$isdir %in% TRUE]
  sort_pid_levels(basename(dd))
}

build_patient_manifest <- function(meta_tbl, base_dir) {
  pid_dirs <- available_pid_dirs(base_dir)
  meta_tbl %>%
    mutate(
      effective_id = ifelse(!is.na(ID_alt) & nzchar(ID_alt), ID_alt, ID),
      pid = as.character(pid),
      Stage = as.character(Stage),
      Age = suppressWarnings(as.numeric(Age))
    ) %>%
    filter(pid %in% pid_dirs) %>%
    group_by(pid) %>%
    group_modify(function(.x, .y) compute_delta_time_days(.x)) %>%
    ungroup() %>%
    mutate(
      rds_path = file.path(base_dir, pid, "perspectives", effective_id, "TranscriptomePerspective.rds"),
      exists = file.exists(rds_path)
    ) %>%
    arrange(factor(pid, levels = sort_pid_levels(pid)), factor(Stage, levels = stage_levels))
}

profile_to_karyotype_strings <- function(profile_mat) {
  if (nrow(profile_mat) != 22) {
    warning("Expected 22 chromosome rows; got ", nrow(profile_mat))
  }
  x <- t(round(as.matrix(profile_mat)))
  apply(x, 1, paste, collapse = ".")
}

build_count_matrix_from_profiles <- function(patient_manifest) {
  stopifnot(length(unique(patient_manifest$pid)) == 1L)
  rows_ok <- patient_manifest[patient_manifest$exists, , drop = FALSE]
  rows_ok <- rows_ok[order(match(rows_ok$Stage, stage_levels)), , drop = FALSE]
  if (nrow(rows_ok) < 2) {
    stop("Need at least two available samples for patient ", unique(patient_manifest$pid))
  }

  long_df <- do.call(rbind, lapply(seq_len(nrow(rows_ok)), function(i) {
    rr <- rows_ok[i, , drop = FALSE]
    obj <- read_transcriptome_perspective(rr$rds_path)
    if (!is.null(obj$passaging_id) && !identical(as.character(obj$passaging_id), as.character(rr$effective_id))) {
      warning("passaging_id mismatch for ", rr$pid, " / ", rr$effective_id)
    }
    tibble(
      karyotype = profile_to_karyotype_strings(obj$profile),
      Delta_Time = rr$Delta_Time,
      Stage = rr$Stage,
      effective_id = rr$effective_id
    )
  }))

  delta_levels <- sort(unique(rows_ok$Delta_Time))
  count_mat <- as.data.frame.matrix(table(
    long_df$karyotype,
    factor(long_df$Delta_Time, levels = delta_levels)
  ))
  colnames(count_mat) <- as.character(delta_levels)
  count_mat <- count_mat[order(rowSums(count_mat), decreasing = TRUE), , drop = FALSE]

  list(
    count_matrix = count_mat,
    long_df = long_df
  )
}

save_alfak_input <- function(count_matrix, out_path) {
  yi <- list(
    x = count_matrix,
    pop.fitness = NULL,
    dt = 1
  )
  saveRDS(yi, out_path)
  invisible(out_path)
}

extract_xval_metrics <- function(xv) {
  metrics <- list(
    xval_r2 = NA_real_,
    xval_cor = NA_real_,
    xval_rmse = NA_real_,
    xval_mae = NA_real_,
    n_xval = NA_integer_
  )

  if (is.null(xv)) {
    return(metrics)
  }

  if (is.atomic(xv) && is.numeric(xv) && is.null(dim(xv)) && length(xv) >= 1L) {
    metrics$xval_r2 <- suppressWarnings(as.numeric(xv[[1]]))
    return(metrics)
  }

  if (!is.list(xv)) {
    return(metrics)
  }

  if (!is.null(xv$R2R)) {
    metrics$xval_r2 <- suppressWarnings(as.numeric(xv$R2R))
  }

  if (is.null(xv$tmp)) {
    return(metrics)
  }

  tmp_df <- as.data.frame(xv$tmp)
  if (ncol(tmp_df) < 2) {
    return(metrics)
  }

  colnames(tmp_df)[1:2] <- c("f_est", "f_xv")
  tmp_df <- tmp_df[, c("f_est", "f_xv"), drop = FALSE]
  tmp_df$f_est <- suppressWarnings(as.numeric(tmp_df$f_est))
  tmp_df$f_xv <- suppressWarnings(as.numeric(tmp_df$f_xv))
  tmp_df <- tmp_df[is.finite(tmp_df$f_est) & is.finite(tmp_df$f_xv), , drop = FALSE]

  metrics$n_xval <- nrow(tmp_df)
  if (!nrow(tmp_df)) {
    return(metrics)
  }

  diff_vec <- tmp_df$f_est - tmp_df$f_xv
  metrics$xval_rmse <- sqrt(mean(diff_vec^2))
  metrics$xval_mae <- mean(abs(diff_vec))
  if (nrow(tmp_df) >= 2) {
    metrics$xval_cor <- suppressWarnings(stats::cor(tmp_df$f_est, tmp_df$f_xv, use = "complete.obs"))
  }
  metrics
}

run_alfak_fit <- function(input_rds,
                          outdir,
                          minobs,
                          pm,
                          nboot,
                          n0,
                          nb,
                          correct_efflux = TRUE,
                          force_refit = FALSE) {
  landscape_path <- file.path(outdir, "landscape.Rds")
  xval_path <- file.path(outdir, "xval.Rds")
  pm_label <- format(pm, scientific = FALSE, trim = TRUE)

  if (!force_refit && file.exists(landscape_path) && file.exists(xval_path)) {
    cached_fit <- tryCatch({
      xv <- readRDS(xval_path)
      c(
        list(
          outdir = outdir,
          pm = pm,
          pm_label = pm_label,
          minobs = minobs,
          status = "ok",
          error_message = NA_character_
        ),
        extract_xval_metrics(xv),
        list(
          landscape_path = landscape_path,
          xval_path = xval_path
        )
      )
    }, error = function(e) NULL)
    if (!is.null(cached_fit)) {
      return(cached_fit)
    }
  }

  dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
  yi <- readRDS(input_rds)
  yi$x <- as.data.frame(yi$x)
  if (diploid_state %in% rownames(yi$x)) {
    yi$x <- yi$x[rownames(yi$x) != diploid_state, , drop = FALSE]
  }
  if (!nrow(yi$x)) {
    stop("All rows were filtered out for input: ", input_rds)
  }
  if (max(rowSums(yi$x), na.rm = TRUE) < minobs) {
    stop("No frequent karyotypes reach minobs=", minobs, " for input: ", input_rds)
  }

  tryCatch({
    alfakR::alfak(
      yi = yi,
      outdir = outdir,
      passage_times = NULL,
      minobs = minobs,
      nboot = nboot,
      n0 = n0,
      nb = nb,
      pm = pm,
      correct_efflux = correct_efflux
    )

    xv <- readRDS(xval_path)
    c(
      list(
        outdir = outdir,
        pm = pm,
        pm_label = pm_label,
        minobs = minobs,
        status = "ok",
        error_message = NA_character_
      ),
      extract_xval_metrics(xv),
      list(
        landscape_path = landscape_path,
        xval_path = xval_path
      )
    )
  }, error = function(e) {
    list(
      outdir = outdir,
      pm = pm,
      pm_label = pm_label,
      minobs = minobs,
      status = "error",
      error_message = conditionMessage(e),
      xval_r2 = NA_real_,
      xval_cor = NA_real_,
      xval_rmse = NA_real_,
      xval_mae = NA_real_,
      n_xval = NA_integer_,
      landscape_path = if (file.exists(landscape_path)) landscape_path else NA_character_,
      xval_path = if (file.exists(xval_path)) xval_path else NA_character_
    )
  })
}

fit_patient_minobs_task <- function(patient_id,
                                    input_rds,
                                    fit_root,
                                    pm_values,
                                    minobs,
                                    nboot,
                                    n0,
                                    nb,
                                    correct_efflux,
                                    force_refit = FALSE) {
  fit_rows <- lapply(pm_values, function(pm_val) {
    pm_label <- format(pm_val, scientific = FALSE, trim = TRUE)
    outdir <- file.path(fit_root, paste0("pm_", pm_label), paste0("MINOBS_", minobs), patient_id)
    run_alfak_fit(
      input_rds = input_rds,
      outdir = outdir,
      minobs = minobs,
      pm = pm_val,
      nboot = nboot,
      n0 = n0,
      nb = nb,
      correct_efflux = correct_efflux,
      force_refit = force_refit
    )
  })

  fit_tbl <- dplyr::bind_rows(lapply(fit_rows, as.data.frame))
  fit_tbl$patient_id <- patient_id
  fit_tbl
}

rank_fit_summary_tbl <- function(fit_tbl) {
  fit_tbl %>%
    dplyr::mutate(
      status = as.character(status),
      rank_xval_r2 = dplyr::coalesce(xval_r2, -Inf),
      rank_xval_cor = dplyr::coalesce(xval_cor, -Inf),
      rank_xval_rmse = dplyr::coalesce(xval_rmse, Inf),
      rank_xval_mae = dplyr::coalesce(xval_mae, Inf)
    ) %>%
    dplyr::arrange(
      factor(patient_id, levels = sort_pid_levels(patient_id)),
      dplyr::desc(status == "ok"),
      dplyr::desc(rank_xval_r2),
      rank_xval_rmse,
      rank_xval_mae,
      dplyr::desc(rank_xval_cor),
      minobs,
      pm
    )
}

select_best_fit_tbl <- function(fit_tbl) {
  ranked_tbl <- rank_fit_summary_tbl(fit_tbl)

  best_tbl <- ranked_tbl %>%
    dplyr::filter(status == "ok") %>%
    dplyr::group_by(patient_id) %>%
    dplyr::slice_head(n = 1) %>%
    dplyr::ungroup() %>%
    dplyr::select(-rank_xval_r2, -rank_xval_cor, -rank_xval_rmse, -rank_xval_mae)

  missing_patients <- setdiff(unique(as.character(ranked_tbl$patient_id)), unique(as.character(best_tbl$patient_id)))
  if (length(missing_patients)) {
    stop("No successful ALFA-K fit found for patient(s): ", paste(missing_patients, collapse = ", "))
  }

  ranked_tbl <- ranked_tbl %>%
    dplyr::select(-rank_xval_r2, -rank_xval_cor, -rank_xval_rmse, -rank_xval_mae)

  list(
    all = ranked_tbl,
    best = best_tbl
  )
}

landscape_to_patient_df <- function(landscape_df, patient_id) {
  k_mat <- do.call(rbind, lapply(landscape_df$k, function(xx) {
    as.numeric(unlist(strsplit(xx, split = "\\.")))
  }))
  colnames(k_mat) <- as.character(seq_len(ncol(k_mat)))
  out <- as.data.frame(k_mat)
  out$karyotypes <- landscape_df$k
  out$fitness <- landscape_df$mean
  out$Nf_01 <- normalize_01(out$fitness)
  out$Nf_mean <- normalize_mean(out$fitness)
  out$Nf_median <- normalize_median(out$fitness)
  out$Patient <- patient_id
  out
}

pairwise_corr <- function(df, value_col) {
  pats <- sort_pid_levels(unique(df$Patient))
  res <- data.frame(
    Patient1 = character(),
    Patient2 = character(),
    Correlation = numeric(),
    Pvalue = numeric(),
    OverlapCount = integer(),
    stringsAsFactors = FALSE
  )
  if (length(pats) < 2) {
    return(res)
  }

  for (i in seq_len(length(pats) - 1L)) {
    for (j in seq.int(i + 1L, length(pats))) {
      p1 <- pats[i]
      p2 <- pats[j]
      d1 <- df[df$Patient == p1, c("karyotypes", value_col), drop = FALSE]
      d2 <- df[df$Patient == p2, c("karyotypes", value_col), drop = FALSE]
      common <- intersect(d1$karyotypes, d2$karyotypes)
      n_overlap <- length(common)
      if (n_overlap > 1) {
        v1 <- d1[[value_col]][match(common, d1$karyotypes)]
        v2 <- d2[[value_col]][match(common, d2$karyotypes)]
        ct <- tryCatch(stats::cor.test(v1, v2, use = "complete.obs"), error = function(e) NULL)
        corr <- if (!is.null(ct)) unname(ct$estimate) else NA_real_
        pval <- if (!is.null(ct)) ct$p.value else NA_real_
      } else {
        corr <- NA_real_
        pval <- NA_real_
      }
      res <- rbind(
        res,
        data.frame(
          Patient1 = p1,
          Patient2 = p2,
          Correlation = corr,
          Pvalue = pval,
          OverlapCount = n_overlap,
          stringsAsFactors = FALSE
        )
      )
    }
  }
  res
}

compute_combined_landscape_metrics <- function(landscape_tables) {
  patient_ids <- sort_pid_levels(names(landscape_tables))
  Data <- bind_rows(lapply(patient_ids, function(pid) landscape_to_patient_df(landscape_tables[[pid]], pid)))

  karyotype_patient <- distinct(Data, Patient, karyotypes)
  nop_tbl <- count(karyotype_patient, karyotypes, name = "Nop")
  Data <- left_join(Data, nop_tbl, by = "karyotypes")

  n_patients <- length(patient_ids)
  near_ubiq_nop <- ceiling(0.8 * n_patients)

  karyotypes_no_dup <- count(nop_tbl, Nop, name = "No. of karyotypes") %>%
    complete(Nop = seq_len(n_patients), fill = list(`No. of karyotypes` = 0L)) %>%
    arrange(Nop)

  Data$NF_diploid <- NA_real_
  for (pid in patient_ids) {
    idx_pid <- which(Data$Patient == pid)
    dip_idx <- idx_pid[Data$karyotypes[idx_pid] == diploid_state]
    if (!length(dip_idx)) {
      next
    }
    mean_dip <- mean(Data$fitness[dip_idx], na.rm = TRUE)
    if (is.finite(mean_dip) && mean_dip != 0) {
      Data$NF_diploid[idx_pid] <- Data$fitness[idx_pid] / mean_dip
    }
  }

  ubiq_karyos <- nop_tbl$karyotypes[nop_tbl$Nop == n_patients]
  Data$NF_ubiquitous_karyotypes <- NA_real_
  corr_mat <- NULL
  if (length(ubiq_karyos) > 0) {
    for (pid in patient_ids) {
      dp <- Data[Data$Patient == pid & Data$karyotypes %in% ubiq_karyos, , drop = FALSE]
      if (!nrow(dp)) {
        next
      }
      mean_pu <- mean(dp$fitness, na.rm = TRUE)
      if (is.finite(mean_pu) && mean_pu != 0) {
        Data$NF_ubiquitous_karyotypes[Data$Patient == pid] <- Data$fitness[Data$Patient == pid] / mean_pu
      }
    }

    corr_input <- matrix(NA_real_, nrow = length(ubiq_karyos), ncol = length(patient_ids))
    rownames(corr_input) <- ubiq_karyos
    colnames(corr_input) <- patient_ids
    for (pid in patient_ids) {
      dp <- Data[Data$Patient == pid & Data$karyotypes %in% ubiq_karyos, c("karyotypes", "NF_ubiquitous_karyotypes"), drop = FALSE]
      dp <- dp[!duplicated(dp$karyotypes), , drop = FALSE]
      corr_input[match(dp$karyotypes, ubiq_karyos), pid] <- dp$NF_ubiquitous_karyotypes
    }
    corr_mat <- suppressWarnings(stats::cor(corr_input, use = "pairwise.complete.obs"))
  }

  results_corr_all <- pairwise_corr(Data, "fitness")
  results_corr_ubiq <- pairwise_corr(Data[!is.na(Data$Nop) & Data$Nop == n_patients, , drop = FALSE], "NF_ubiquitous_karyotypes")
  results_corr_near <- pairwise_corr(Data[!is.na(Data$Nop) & Data$Nop >= near_ubiq_nop, , drop = FALSE], "NF_ubiquitous_karyotypes")

  list(
    Data = Data,
    patient_ids = patient_ids,
    n_patients = n_patients,
    karyotypes_no_dup = karyotypes_no_dup,
    results_corr_all = results_corr_all,
    results_corr_ubiq = results_corr_ubiq,
    results_corr_near = results_corr_near,
    corr_mat = corr_mat
  )
}

beneficial_cnv_profile <- function(landscape_df) {
  fq_flag <- landscape_df$fq
  fq_keep <- if (is.logical(fq_flag)) {
    fq_flag
  } else {
    as.character(fq_flag) %in% c("TRUE", "T", "true", "1")
  }
  nn_flag <- landscape_df$nn
  nn_keep <- if (is.logical(nn_flag)) {
    nn_flag
  } else {
    as.character(nn_flag) %in% c("TRUE", "T", "true", "1")
  }

  k_vec <- as.character(landscape_df$k)
  move_labels <- as.vector(rbind(paste0(seq_len(22), "+"), paste0(seq_len(22), "-")))
  prop <- rep(NA_real_, length(move_labels))
  valid_n <- integer(length(move_labels))
  beneficial_n <- integer(length(move_labels))
  names(prop) <- move_labels
  names(valid_n) <- move_labels
  names(beneficial_n) <- move_labels

  if (!length(k_vec)) {
    return(list(
      proportion = prop,
      beneficial_n = beneficial_n,
      valid_n = valid_n
    ))
  }

  k_parts <- strsplit(k_vec, ".", fixed = TRUE)
  part_lengths <- lengths(k_parts)
  if (any(part_lengths != 22L)) {
    bad_idx <- which(part_lengths != 22L)[1]
    stop("Found karyotype with != 22 entries in beneficial_cnv_profile(): ", k_vec[bad_idx])
  }
  k_mat <- matrix(as.integer(unlist(k_parts, use.names = FALSE)), ncol = 22, byrow = TRUE)
  fitness_vec <- as.numeric(landscape_df$mean)

  starts <- which(fq_keep %in% TRUE)
  if (!length(starts)) {
    return(list(
      proportion = prop,
      beneficial_n = beneficial_n,
      valid_n = valid_n
    ))
  }

  edge_rows <- vector("list", length(starts))
  edge_idx <- 0L
  for (from in starts) {
    v <- k_mat[from, ]
    neigh_mat <- matrix(rep(v, each = 44L), nrow = 44L, ncol = 22)
    for (chr in seq_len(22)) {
      plus_row <- 2L * chr - 1L
      minus_row <- 2L * chr
      neigh_mat[plus_row, chr] <- neigh_mat[plus_row, chr] + 1L
      neigh_mat[minus_row, chr] <- neigh_mat[minus_row, chr] - 1L
    }

    neigh_str <- do.call(paste, c(as.data.frame(neigh_mat), sep = "."))
    to <- match(neigh_str, k_vec)
    valid <- !is.na(to)
    if (any(valid)) {
      valid[valid] <- nn_keep[to[valid]] %in% TRUE
    }
    if (!any(valid)) {
      next
    }

    to_valid <- to[valid]
    move_valid <- move_labels[valid]
    delta <- fitness_vec[to_valid] - fitness_vec[from]
    ok <- is.finite(delta)
    benefit <- rep(NA, length(delta))
    benefit[ok] <- delta[ok] > 0

    edge_idx <- edge_idx + 1L
    edge_rows[[edge_idx]] <- data.frame(
      move = move_valid,
      ok = ok,
      benefit = benefit,
      stringsAsFactors = FALSE
    )
  }

  if (edge_idx > 0L) {
    edges_df <- dplyr::bind_rows(edge_rows[seq_len(edge_idx)])
    for (mv in move_labels) {
      ii <- which(edges_df$move == mv)
      if (!length(ii)) {
        next
      }
      sub <- edges_df[ii, , drop = FALSE]
      ok <- sub$ok %in% TRUE
      valid_n[mv] <- sum(ok)
      if (!valid_n[mv]) {
        next
      }
      beneficial_n[mv] <- sum(sub$benefit[ok] %in% TRUE, na.rm = TRUE)
      prop[mv] <- beneficial_n[mv] / valid_n[mv]
    }
  }

  list(
    proportion = prop,
    beneficial_n = beneficial_n,
    valid_n = valid_n
  )
}

parse_segment_table <- function(segment_names) {
  tibble(
    segment = segment_names,
    chromosome = sub(":.*$", "", segment_names),
    start = suppressWarnings(as.numeric(sub("^[^:]+:(\\d+)-.*$", "\\1", segment_names))),
    end = suppressWarnings(as.numeric(sub("^.*-(\\d+)$", "\\1", segment_names)))
  ) %>%
    mutate(
      chromosome = suppressWarnings(as.integer(chromosome)),
      seg_length = ifelse(is.finite(start) & is.finite(end), end - start + 1, NA_real_)
    )
}

build_chromosome_cn_from_profiles <- function(profile_objects) {
  mat_list <- lapply(names(profile_objects), function(sample_id) {
    obj <- profile_objects[[sample_id]]
    x <- as.matrix(obj$profile)
    storage.mode(x) <- "numeric"
    colnames(x) <- paste0(sample_id, "__", colnames(x))
    x
  })
  names(mat_list) <- names(profile_objects)
  seg_mat <- do.call(cbind, mat_list)
  seg_tbl <- parse_segment_table(rownames(seg_mat))
  keep_idx <- which(seg_tbl$chromosome %in% seq_len(22))
  seg_mat <- seg_mat[keep_idx, , drop = FALSE]
  seg_tbl <- seg_tbl[keep_idx, , drop = FALSE]

  selected_idx <- integer(0)
  for (chr in seq_len(22)) {
    idx_chr <- which(seg_tbl$chromosome == chr)
    if (!length(idx_chr)) {
      next
    }
    if (length(idx_chr) == 1L) {
      selected_idx <- c(selected_idx, idx_chr)
    } else {
      lengths_chr <- seg_tbl$seg_length[idx_chr]
      lengths_chr[!is.finite(lengths_chr)] <- -Inf
      selected_idx <- c(selected_idx, idx_chr[which.max(lengths_chr)])
    }
  }

  out <- matrix(
    2,
    nrow = ncol(seg_mat),
    ncol = 22,
    dimnames = list(colnames(seg_mat), as.character(seq_len(22)))
  )
  if (length(selected_idx)) {
    sel_chr <- as.character(seg_tbl$chromosome[selected_idx])
    out[, sel_chr] <- t(seg_mat[selected_idx, , drop = FALSE])
  }
  out
}

build_timepoint_by_cell <- function(patient_manifest, profile_objects) {
  vec_list <- lapply(seq_len(nrow(patient_manifest)), function(i) {
    rr <- patient_manifest[i, , drop = FALSE]
    obj <- profile_objects[[rr$effective_id]]
    setNames(
      rep(as.character(rr$Stage), ncol(obj$profile)),
      paste0(rr$effective_id, "__", colnames(obj$profile))
    )
  })
  unlist(vec_list)
}

save_ploidy_center_heatmap <- function(chromosome_cn,
                                       timepoint_by_cell,
                                       output_pdf,
                                       timepoint_levels = stage_levels,
                                       clustering_method = "ward.D",
                                       pdf_size = 8) {
  heatmap_mat <- as.matrix(chromosome_cn)
  storage.mode(heatmap_mat) <- "numeric"

  timepoint_by_cell <- timepoint_by_cell[rownames(heatmap_mat)]
  timepoint_factor <- factor(as.character(timepoint_by_cell), levels = timepoint_levels)
  timepoint_factor <- droplevels(timepoint_factor)
  timepoint_display <- factor(
    dplyr::recode(as.character(timepoint_factor), Primary = "t9192", Recurrent = "t9907"),
    levels = c("t9192", "t9907")
  )
  timepoint_palette <- c(t9192 = "#33A02C", t9907 = "#FB9A99")

  left_ha <- ComplexHeatmap::rowAnnotation(
    TimePoint = timepoint_display,
    col = list(TimePoint = timepoint_palette),
    simple_anno_size = grid::unit(3.5, "mm"),
    annotation_name_side = "top",
    annotation_name_gp = grid::gpar(fontsize = 9)
  )

  row_clust <- if (nrow(heatmap_mat) > 1) {
    stats::as.dendrogram(stats::hclust(stats::dist(heatmap_mat), method = clustering_method))
  } else {
    FALSE
  }
  col_clust <- if (ncol(heatmap_mat) > 1) {
    cd <- stats::as.dendrogram(stats::hclust(stats::dist(t(heatmap_mat)), method = clustering_method))
    burden <- colSums(heatmap_mat != 2, na.rm = TRUE)
    cd <- stats::reorder(cd, wts = burden, agglo.FUN = mean)
    rev(cd)
  } else {
    FALSE
  }

  cnv_colors <- circlize::colorRamp2(c(0, 2, 4), c("#2166AC", "#E6E6E6", "#B2182B"))
  ht <- ComplexHeatmap::Heatmap(
    heatmap_mat,
    name = "Copy number",
    col = cnv_colors,
    cluster_rows = row_clust,
    cluster_columns = col_clust,
    row_dend_reorder = FALSE,
    column_dend_reorder = FALSE,
    show_column_names = TRUE,
    show_row_names = FALSE,
    left_annotation = left_ha,
    use_raster = TRUE,
    raster_quality = 2,
    row_title = "Cells",
    column_title = NULL,
    heatmap_legend_param = list(at = c(0, 2, 4), labels = c("0", "2", "4"))
  )

  grDevices::pdf(output_pdf, width = pdf_size, height = pdf_size, useDingbats = FALSE)
  on.exit(grDevices::dev.off(), add = TRUE)
  ComplexHeatmap::draw(
    ht,
    merge_legends = TRUE,
    heatmap_legend_side = "right",
    annotation_legend_side = "right"
  )
  invisible(ht)
}

make_focus_umap_plot <- function(landscape_df, patient_id) {
  k_df <- landscape_to_patient_df(landscape_df, patient_id)
  dip_idx <- which(k_df$karyotypes == diploid_state)
  if (!length(dip_idx)) {
    stop("No diploid karyotype found for patient ", patient_id)
  }
  dip_mean <- mean(k_df$fitness[dip_idx], na.rm = TRUE)
  if (!is.finite(dip_mean) || dip_mean == 0) {
    stop("Diploid mean fitness is invalid for patient ", patient_id)
  }

  set.seed(3.1415926)
  umap_mat <- uwot::umap2(as.matrix(k_df[, as.character(seq_len(22))]), n_components = 2, min_dist = 0.9)
  umap_df <- as.data.frame(umap_mat)
  colnames(umap_df) <- c("UMAP1", "UMAP2")
  umap_df$relative_fitness <- k_df$fitness / dip_mean
  umap_df <- umap_df[order(umap_df$relative_fitness, decreasing = FALSE, na.last = TRUE), , drop = FALSE]

  min_rf <- min(umap_df$relative_fitness, na.rm = TRUE)
  color_limits <- c(min(min_rf, 1), 1.5)
  color_limits <- ensure_center_in_limits(color_limits, center = 1)
  if (diff(color_limits) == 0) {
    color_limits[2] <- color_limits[2] + 1e-06
  }
  color_knots <- c(
    seq(color_limits[1], 1, length.out = 6),
    seq(1, color_limits[2], length.out = 6)[-1]
  )
  color_values <- scales::rescale(color_knots, from = color_limits)

  ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = relative_fitness)) +
    geom_jitter(width = 0.1, height = 0.1, alpha = 1, size = 1, shape = 16) +
    scale_color_gradientn(
      colours = rev(RColorBrewer::brewer.pal(11, "RdYlBu")),
      values = color_values,
      limits = color_limits,
      breaks = c(color_limits[1], 1, color_limits[2]),
      labels = signif(c(color_limits[1], 1, color_limits[2]), 3),
      oob = squish,
      name = "Relative fitness\n(vs diploid)"
    ) +
    theme_minimal()
}
```

## Metadata and Available Patients

```{r metadata}
meta_tbl <- readxl::read_excel(meta_path)
patient_manifest <- build_patient_manifest(meta_tbl, data_dir)

patient_summary <- patient_manifest %>%
  group_by(pid) %>%
  summarise(
    n_samples_in_meta = n(),
    n_available = sum(exists),
    stages_available = paste(sort(unique(Stage[exists])), collapse = ", "),
    effective_ids = paste(effective_id[exists], collapse = ", "),
    recurrent_delta_days = unique(Delta_Time[Stage == "Recurrent"])[1],
    .groups = "drop"
  ) %>%
  arrange(factor(pid, levels = sort_pid_levels(pid)))

readr::write_tsv(patient_manifest, file.path(result_dir, "patient_manifest.tsv"))
readr::write_tsv(patient_summary, file.path(result_dir, "patient_summary.tsv"))

patient_summary
```

```{r fitting-patients}
fit_manifest <- patient_manifest %>%
  filter(exists) %>%
  group_by(pid) %>%
  filter(all(stage_levels %in% Stage)) %>%
  ungroup()

fit_patient_ids <- sort_pid_levels(unique(fit_manifest$pid))
fit_patient_ids
```

## Build ALFA-K Inputs

```{r build-inputs}
input_index <- lapply(fit_patient_ids, function(pid) {
  df_pid <- fit_manifest %>% filter(pid == !!pid)
  obj <- build_count_matrix_from_profiles(df_pid)
  out_path <- file.path(input_dir, paste0(pid, ".Rds"))
  save_alfak_input(obj$count_matrix, out_path)
  readr::write_tsv(
    tibble(karyotype = rownames(obj$count_matrix)) %>%
      bind_cols(as.data.frame(obj$count_matrix)),
    file.path(result_dir, paste0(pid, "_ALFAK_input_counts.tsv"))
  )
  tibble(
    pid = pid,
    input_rds = out_path,
    n_karyotypes = nrow(obj$count_matrix),
    timepoints = paste(colnames(obj$count_matrix), collapse = ", ")
  )
})
input_index_tbl <- bind_rows(input_index)
readr::write_tsv(input_index_tbl, file.path(result_dir, "alfak_input_index.tsv"))
input_index_tbl
```

## Refit Landscapes with Current ALFA-K

```{r refit-landscapes}
if (!requireNamespace("foreach", quietly = TRUE) || !requireNamespace("doParallel", quietly = TRUE)) {
  stop("Packages 'foreach' and 'doParallel' are required for parallel ALFA-K fitting.")
}

fit_summary_path <- file.path(result_dir, "fit_summary.tsv")
selected_fit_path <- file.path(result_dir, "selected_fit.tsv")

if (!force_refit && file.exists(fit_summary_path) && file.exists(selected_fit_path)) {
  message("Reusing existing fit_summary.tsv and selected_fit.tsv; skipping ALFA-K task dispatch.")
  fit_summary_tbl <- readr::read_tsv(fit_summary_path, show_col_types = FALSE)
  selected_fit_tbl <- readr::read_tsv(selected_fit_path, show_col_types = FALSE)
} else {
  fit_tasks_tbl <- expand.grid(
    patient_id = fit_patient_ids,
    minobs = minobs_values_use,
    stringsAsFactors = FALSE
  )
  fit_tasks_tbl$input_rds <- file.path(input_dir, paste0(fit_tasks_tbl$patient_id, ".Rds"))
  fit_tasks_tbl <- fit_tasks_tbl[file.exists(fit_tasks_tbl$input_rds), , drop = FALSE]
  fit_tasks_tbl <- fit_tasks_tbl %>%
    dplyr::arrange(factor(patient_id, levels = fit_patient_ids), minobs)

  if (!nrow(fit_tasks_tbl)) {
    stop("No ALFA-K fitting tasks were constructed.")
  }

  n_cores_fit <- max(1L, min(n_cores_use, nrow(fit_tasks_tbl)))
  message("Dispatching ", nrow(fit_tasks_tbl), " ALFA-K tasks across ", n_cores_fit, " core(s).")

  readr::write_tsv(fit_tasks_tbl, file.path(result_dir, "alfak_task_index.tsv"))

  if (n_cores_fit == 1L) {
    fit_task_results <- lapply(seq_len(nrow(fit_tasks_tbl)), function(i) {
      rr <- fit_tasks_tbl[i, , drop = FALSE]
      fit_patient_minobs_task(
        patient_id = rr$patient_id,
        input_rds = rr$input_rds,
        fit_root = fit_dir,
        pm_values = pm_values,
        minobs = rr$minobs,
        nboot = nboot_use,
        n0 = n0_use,
        nb = nb_use,
        correct_efflux = correct_efflux_use,
        force_refit = force_refit
      )
    })
  } else {
    cl <- parallel::makeCluster(n_cores_fit)
    parallel::clusterSetRNGStream(cl, iseed = 31415926)
    doParallel::registerDoParallel(cl)
    fit_task_results <- tryCatch({
      foreach::foreach(
        i = seq_len(nrow(fit_tasks_tbl)),
        .packages = c("alfakR", "dplyr")
      ) %dopar% {
        rr <- fit_tasks_tbl[i, , drop = FALSE]
        fit_patient_minobs_task(
          patient_id = rr$patient_id,
          input_rds = rr$input_rds,
          fit_root = fit_dir,
          pm_values = pm_values,
          minobs = rr$minobs,
          nboot = nboot_use,
          n0 = n0_use,
          nb = nb_use,
          correct_efflux = correct_efflux_use,
          force_refit = force_refit
        )
      }
    }, finally = {
      try(parallel::stopCluster(cl), silent = TRUE)
      foreach::registerDoSEQ()
    })
  }

  saveRDS(fit_task_results, file.path(result_dir, "fit_task_results.Rds"))

  fit_summary_tbl <- dplyr::bind_rows(fit_task_results)
  fit_selection <- select_best_fit_tbl(fit_summary_tbl)
  fit_summary_tbl <- fit_selection$all
  selected_fit_tbl <- fit_selection$best

  readr::write_tsv(fit_summary_tbl, fit_summary_path)
  readr::write_tsv(selected_fit_tbl, selected_fit_path)
}

if (nrow(selected_fit_tbl) != length(fit_patient_ids)) {
  stop("Expected one selected fit per patient; got ", nrow(selected_fit_tbl), " rows for ", length(fit_patient_ids), " patients.")
}

selected_fit_tbl %>%
  arrange(factor(patient_id, levels = sort_pid_levels(patient_id)))
```

## Load Selected Landscapes

```{r load-landscapes}
selected_landscapes <- lapply(seq_len(nrow(selected_fit_tbl)), function(i) {
  rr <- selected_fit_tbl[i, , drop = FALSE]
  x <- readRDS(rr$landscape_path)
  x
})
names(selected_landscapes) <- selected_fit_tbl$patient_id

combined_metrics <- compute_combined_landscape_metrics(selected_landscapes)
saveRDS(combined_metrics, file.path(result_dir, "combined_metrics.rds"))

combined_metrics$karyotypes_no_dup
```

## Panel B

```{r panel-b}
if (!focus_pid %in% fit_patient_ids) {
  stop("focus_pid ", focus_pid, " is not available with both Primary and Recurrent samples.")
}

focus_manifest <- fit_manifest %>%
  filter(pid == focus_pid) %>%
  arrange(factor(Stage, levels = stage_levels))

focus_profiles <- lapply(seq_len(nrow(focus_manifest)), function(i) {
  rr <- focus_manifest[i, , drop = FALSE]
  obj <- read_transcriptome_perspective(rr$rds_path)
  obj
})
names(focus_profiles) <- focus_manifest$effective_id

focus_chrom_cn <- build_chromosome_cn_from_profiles(focus_profiles)
focus_timepoint_by_cell <- build_timepoint_by_cell(focus_manifest, focus_profiles)

panel_b_pdf <- file.path(figure_dir, paste0("Panel_B_", focus_pid, "_ploidy_center.pdf"))
panel_b_ht <- save_ploidy_center_heatmap(
  chromosome_cn = focus_chrom_cn,
  timepoint_by_cell = focus_timepoint_by_cell,
  output_pdf = panel_b_pdf,
  timepoint_levels = stage_levels,
  clustering_method = "ward.D",
  pdf_size = 8
)
ComplexHeatmap::draw(
  panel_b_ht,
  merge_legends = TRUE,
  heatmap_legend_side = "right",
  annotation_legend_side = "right"
)

cat("Saved Panel B to:", panel_b_pdf)
```

## Panel C

```{r panel-c}
focus_landscape <- selected_landscapes[[focus_pid]]
panel_c_plot <- make_focus_umap_plot(focus_landscape, focus_pid)
panel_c_pdf <- file.path(figure_dir, paste0("Panel_C_", focus_pid, "_relative_diploid_umap.pdf"))
save_plot_pdf(panel_c_plot, panel_c_pdf, width = 5.5, height = 5)
panel_c_plot
```

## Panel D

```{r panel-d}
beneficial_profiles <- lapply(names(selected_landscapes), function(pid) {
  beneficial_cnv_profile(selected_landscapes[[pid]])
})
names(beneficial_profiles) <- names(selected_landscapes)

beneficial_list <- lapply(beneficial_profiles, function(x) x$proportion)
beneficial_valid_n_list <- lapply(beneficial_profiles, function(x) x$valid_n)

beneficial_mat <- do.call(rbind, beneficial_list)
beneficial_mat <- beneficial_mat[sort_pid_levels(rownames(beneficial_mat)), , drop = FALSE]
readr::write_tsv(
  tibble::rownames_to_column(as.data.frame(beneficial_mat), var = "Patient"),
  file.path(result_dir, "beneficial_karyotype_proportion_matrix.tsv")
)

beneficial_valid_n_mat <- do.call(rbind, beneficial_valid_n_list)
beneficial_valid_n_mat <- beneficial_valid_n_mat[sort_pid_levels(rownames(beneficial_valid_n_mat)), , drop = FALSE]
readr::write_tsv(
  tibble::rownames_to_column(as.data.frame(beneficial_valid_n_mat), var = "Patient"),
  file.path(result_dir, "beneficial_karyotype_valid_move_count_matrix.tsv")
)

panel_d_pdf <- file.path(figure_dir, "Panel_D_beneficial_karyotype_proportion_heatmap.pdf")
col_fun_d <- circlize::colorRamp2(c(0, 0.5, 1), c("blue", "#EEEEEE", "red"))
panel_d_mat <- beneficial_mat
panel_d_mat[!is.finite(panel_d_mat)] <- 0

ht_d <- ComplexHeatmap::Heatmap(
  panel_d_mat,
  name = "Beneficial\nkaryotypes\nproportion",
  col = col_fun_d,
  cluster_rows = nrow(panel_d_mat) > 1,
  cluster_columns = ncol(panel_d_mat) > 1,
  na_col = "#BDBDBD",
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_names_gp = grid::gpar(fontsize = 8),
  column_names_gp = grid::gpar(fontsize = 7),
  rect_gp = grid::gpar(col = "white", lwd = 1)
)

grDevices::pdf(panel_d_pdf, width = 10, height = 4.5, useDingbats = FALSE)
ComplexHeatmap::draw(ht_d, heatmap_legend_side = "right")
grDevices::dev.off()
ComplexHeatmap::draw(ht_d, heatmap_legend_side = "right")

cat("Saved Panel D to:", panel_d_pdf)
```

## Panel E

```{r panel-e}
panel_e_df <- combined_metrics$karyotypes_no_dup %>%
  mutate(Nop = factor(Nop, levels = seq_len(combined_metrics$n_patients)))

panel_e_plot <- ggplot(panel_e_df, aes(x = Nop, y = `No. of karyotypes`, fill = Nop)) +
  geom_col(width = 0.8) +
  geom_text(aes(label = `No. of karyotypes`), vjust = -0.4, size = 3) +
  scale_fill_manual(values = grDevices::colorRampPalette(c("#7AD151", "#22A884", "#FCAE61", "#D73027"))(nrow(panel_e_df))) +
  labs(
    x = "Number of Patients",
    y = "Number of Karyotypes"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

panel_e_pdf <- file.path(figure_dir, "Panel_E_karyotypes_by_patient_overlap.pdf")
save_plot_pdf(panel_e_plot, panel_e_pdf, width = 5.5, height = 4.5)
panel_e_plot
```

## Panel F

```{r panel-f}
shared_sig <- combined_metrics$results_corr_all %>%
  filter(!is.na(Pvalue), !is.na(Correlation), Pvalue < 0.05, Correlation > 0) %>%
  mutate(label = "shared")

ubiq_sig <- combined_metrics$results_corr_ubiq %>%
  filter(!is.na(Pvalue), !is.na(Correlation), Pvalue < 0.05, Correlation > 0) %>%
  mutate(label = "ubiquitous")

panel_f_df <- bind_rows(shared_sig, ubiq_sig) %>%
  mutate(label = factor(label, levels = c("shared", "ubiquitous")))

panel_f_plot <- if (length(unique(na.omit(panel_f_df$label))) >= 2 && nrow(panel_f_df) > 1) {
  ggpubr::ggviolin(
    panel_f_df,
    x = "label",
    y = "Correlation",
    fill = "label",
    add = "boxplot",
    palette = c(shared = "#1F77B4", ubiquitous = "#FF7F0E")
  ) +
    ggpubr::stat_compare_means(
      comparisons = list(c("shared", "ubiquitous")),
      label = "p.format",
      method = "t.test"
    ) +
    ggpubr::rremove("legend") +
    labs(x = "Karyotype Type", y = "Pearson Correlation") +
    theme_minimal()
} else {
  placeholder_plot("Insufficient shared/ubiquitous correlation data")
}

panel_f_pdf <- file.path(figure_dir, "Panel_F_shared_vs_ubiquitous_correlation_violin.pdf")
save_plot_pdf(panel_f_plot, panel_f_pdf, width = 4.5, height = 4.5)
panel_f_plot
```

## Panel G

```{r panel-g}
panel_g_df <- combined_metrics$results_corr_all %>%
  mutate(
    Patient1 = factor(Patient1, levels = combined_metrics$patient_ids),
    Patient2 = factor(Patient2, levels = combined_metrics$patient_ids),
    label = ifelse(
      is.na(Pvalue),
      "",
      ifelse(Pvalue < 0.001, "***",
             ifelse(Pvalue < 0.01, "**",
                    ifelse(Pvalue < 0.05, "*", "")))
    )
  )

panel_g_plot <- ggplot(panel_g_df, aes(x = Patient1, y = Patient2)) +
  geom_point(aes(size = OverlapCount, color = Correlation)) +
  geom_text(aes(label = label), vjust = -1, size = 3) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, na.value = "grey80") +
  labs(
    title = "Pairwise Correlation of Shared Karyotype Fitness",
    x = "",
    y = "",
    size = "Shared Karyotype\nOverlap Count",
    color = "Correlation"
  ) +
  theme_minimal()

panel_g_pdf <- file.path(figure_dir, "Panel_G_pairwise_shared_karyotype_correlation.pdf")
save_plot_pdf(panel_g_plot, panel_g_pdf, width = 7, height = 6)
panel_g_plot
```

## Output Files

```{r output-summary}
output_files <- tibble(
  panel = c("B", "C", "D", "E", "F", "G"),
  file = c(
    panel_b_pdf,
    panel_c_pdf,
    panel_d_pdf,
    panel_e_pdf,
    panel_f_pdf,
    panel_g_pdf
  )
)

output_files
```

## Session Info

```{r session-info}
sessionInfo()
```
